---
title: "ECON03GE3A"
subtitle: "The Colonial Legacy"
author: ""
csl: harvard-cite-them-right.csl
bibliography: ref.bib
date: "2021-08-20"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Reference 

The Colonial Legacy 

Chandra (2004)

![](https://books.google.com/books/publisher/content/images/frontcover/DgESBQAAQBAJ?fife=w200-h300)



## The Colonial Structure {.smaller }

The four basic features of the colonial structure in India:

- Complete but complex integration and enmeshing of the colony with the world capitalist system in a subordinate position.

- Unequal exchange and internal disarticulation of the economy [India's Foreign Trade](1.TheColonialLegacy/Table1.PNG)

- Availability and utilisation of surpluses for development was low. [Capital Formation and Savings 1901-46](1.TheColonialLegacy/Table2.PNG)
  - Why was the availability of surplus low?
    - Drain
    - Appropriation by the colonial state
    - Landlords and moneylenders
      
- Policies of the colonial state in the subordination of India.


## Colonial Impact on Agriculture {.smaller }

- Severe impact during the first half of the twentieth century.
  - Per capita agricultural production declined by 0.72% per year during 1911-14.
  - Per capita food grain output declined by 1.14% per year.

- Colonialism made the agrarian structure extremely regressive.
  - *Zamindars* failed to invest in land.
  - More reliance on rack-renting.
  - Peasants lost land to the moneylenders.
  - Sub infeudation, tenancy, and sharecropping dominated the *zamindari* and the *ryotwari* areas.
  
  
## {.smaller}


- No change in the technology or techniques and inputs.
```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(tidyverse); library(kableExtra)
tibble(
  `Technology` = c('No. of ploughs used in India (late 1930s)',
         'Agricultural department selling improved ploughs', 'Use of improved seeds (1922-23)',
         'Use of chemical fertilizers ',
         'Number of agricultural colleges in 1946',
         "Investment in terracing, flood control,\ndrainage and de-salinization of soil",
         '% of total cultivated area irrigated'
         ), 
  `Statistics` = c('32 million','7 to 8 thousand per year','1.9%',
                   '2000 tonnes per year (1898-1923)\n17,400 tonnes per year (1919-24)',
                   '9 with 3110 students',
                   'Nil',
                   '26.7% in 1940'
                   )
) %>% kable() %>% 
  kable_styling(bootstrap_options = c("hover", "striped"))
  
```


## Colonial impact on Industry

- Decline in urban handicrafts
- Slow growth of modern industries
  - Confined to jute and textile
  - Iron and steel after 1907
  - Sugar, cement, and paper after 1930s
    - Depression induced weaknesses
- Cotton and jute accounted for 30% of workforce until 1946.
- Growth rate: 3.8% per annum
- Share of industry: 7.5% in 1918, 3.8% in 1913


## Colonial impact on Industry

- Employment:
  - In 1939, Population = 389 million, employed  = 2 million
  - In 1951 - 2.3% employed in industry
- [The composition/structure of India's national product](1.TheColonialLegacy/Table3.PNG)   
- Capital goods industry
  - Absence of capital goods industry.
  - Relied mostly on imported machinery or tools.
  - 89.9% of capital goods imported in 1950.
  
  
## Colonial impact on industry {.smaller}

- Banking Sector/industry
  - Underdeveloped
```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(tidyverse); library(kableExtra)
tibble(
  Year = c(1914, 1940, 2009
         ), 
  Statistics = c('1 bank office for 1.7 million people',
                 '1 bank office for 90,000 people',
                 '1 bank office for 14 people'
                   )
) %>% kable() %>% 
  kable_styling(bootstrap_options = c("hover", "striped"))
  
```

- Trade
  - Imports $\rightarrow$ indigenous industries, modern industries.
  - Displaced indigenous industries.





## Colonial impact on industry


- Railways
  - Commercial revolution $\rightarrow$ colonialisation $\rightarrow$ exports and distribution.
  
- Foreign investment
  - Uniliver, Imperial Chemical Industries, Dunlop, and General Motors entry after 1918.
  - Contributed to 'guided underdevelopment'.
  - Production and export of raw materials and food stuffs.
  - Catered to foreign markets.
  

## Other economic indicators of backwardness

- Very slow growth in India's national income and stagnation in per capita income [National Product](1.TheColonialLegacy/Table4.PNG)
  - Angus Maddison: [The Maddison Project](https://www.rug.nl/ggdc/historicaldevelopment/maddison/releases/maddison-project-database-2020)
  - Alan Heston: [Penn World Table](https://www.rug.nl/ggdc/productivity/pwt/)


## Other economic indicators of backwardness {.smaller}

```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(tidyverse); library(kableExtra)
tibble(
  Year = c('Cereals and Pulses availability (1950)',
           'Availability of cloth (1950)',
           'Death rate (1950)',
           'Infant Mortality Rate (1950)',
           'Life expectancy (1950)',
           'No of medical colleges (1943)',
           'No of medical doctors (1951)',
           'Cases of small pox, plague, and cholera deaths',
           '% of illiterate population (1947) '
         ), 
  Statistics = c('394.9 grams/day/person',
                 '10 meters/year/person',
                 '27.4 / 1000 persons',
                 '175 - 190 / 1000 live births',
                 '32 years',
                 '10 with 700 graduates',
                 '18000',
                 'in millions',
                 '88%'
                   )
) %>% kable() %>% 
  kable_styling(bootstrap_options = c("hover", "striped"))
  
```


## Summary::India's economic profile in 1940s

- Stagnating per capita national income.
- Abysmal standard of living.
- Stunted industrial development
- Stagnant, low-productivity, semi-fudal agriculture.

